---
title: "Final Project Draft"
author: "Lindsey Fisher"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
```

## An Analysis of US Enenrgy Related Consumption and Emmisions Data
### Group Members: Lindsey Fisher, Param Bhandare, Yunqing Xiao

```{r, include=FALSE}

E2020 = read_csv("emissions2020.csv")
C2020 = read_csv("consumption2020.csv")


E2020A = E2020 %>% 
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE)
    
E2020B = E2020A %>%    
mutate(NumYear = as.numeric(as.character(unlist(E2020A$Year)))) %>% 
filter(Sector != "NA")

    

C2020A = C2020 %>%
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2020B = C2020A %>% 
mutate(YearN = as.numeric(as.character(unlist(C2020A$Year))),
       RateNumeric = as.numeric(sub("%","",C2020A$`Growth (2019-2050)`))) 
    

```


```{r}
E2011 =read_csv("emissions2011.csv")
C2011 = read_csv("consumption2011.csv")

E2011 = E2011 %>% 
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE) 
    
E2011A = E2011 %>%    
select(Year, Sector, Source, MmtCO2e,"Growth Rate (2009-2035)") %>% 
    filter(Sector != "NA")


C2011A = C2011 %>%
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2011A = C2011A %>% 
select(Year, Sector, Source, quads,"Growth Rate (2009-2035)") %>% 
 filter(Source != "Total") %>% 
    arrange(Year)

```


### Introduction

We were interested in looking at something environmentally related, so we chose US energy consumption and emissions. We chose two reports from the Annual Energy Outlook (AEO) report from the US Energy Information Administration. We chose to look at the AEO published in 2011, and the AEO published in 2020. We chose these reports because they are a decade apart and will serve as a comparison to see if we have reduced emissions and consumption overall. We believe there is a lot of relationships between variables to be examined and uncovered. We are interested in relationships between the variables in each of the data sets as well as comparing the variables across the separate data sets to look for trends. We hope to use this analysis to locate key areas where the US is doing well in reducing emissions and consumption and areas where we need to focus on improvement. We hope to assess the validity of the AEOs projections by comparing the future projections in the 2011 report to the observed measures from the 2020 report. 

### Background

Emissions Data: 
The emissions data measures outputs into the atmosphere in Million Metric Ton CO2 Equivalents. This is a measure of global warming potential. CO2 is a well know gas that traps heat in the atmosphere, however, there are other gasses which trap heat at a much higher magnitude. The measurement is called CO2 equivalents because it measures the amount of total global warming potential as a factor of the magnitude of one CO2 molecules warming potential. It is separated out by sector: residential, commercial, industrial, transportation, and electric power. As well as by fuel source: petroleum, natural gas, coal, and electric power generation, and an “other” category.  The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 

Consumption Data: 
The consumption data measures how much energy was used in units of Quads. A Quad is a unit of energy equal to 10¹⁵ British Thermal Units, or 1.055 × 10¹⁸ joules in SI units. The US Department of Energy adopted the use of quads as their unit of energy use measurement for national and international reports.  The consumption data reports energy consumption by sector and power source. Sectors include residential, commercial, industrial, transportation, electric power, and delivered energy. The energy sources for consumption are more complex and vary with in each sector. The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 
    
### Analysis


#### 1. How does total emissions vary by each year?Which source is the greatest contributer for each year?

```{r}
totalE = E2020B %>% 
filter(Source == "Total") %>% 
arrange(NumYear)

ggplot(totalE,aes(NumYear,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")

sourcetopn = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(NumYear) %>% 
    slice_max(order_by = MmtCO2e, n = 3)

ggplot(sourcetopn, aes(NumYear, MmtCO2e, color = Source))+
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")
```

```{r}

total_2011 = E2011A %>% 
    filter(!str_detect(Source, "Total"))%>%
    arrange(Year)
    

ggplot(total_2011,aes(Year,MmtCO2e))+
    stat_smooth()+
    geom_point()

sourcetopn_2011 = E2011A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    group_by(Year) %>% 
    slice_max(order_by = MmtCO2e, n = 3)

ggplot(sourcetopn_2011, aes(Year, MmtCO2e, color = Source))+
    geom_point()

```

#### 2.Which sector produces the most emissions? Does this vary by year?

```{r}
sectormostyear = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(NumYear, Sector) %>% 
    top_n(MmtCO2e, n = 1)

sectormost = E2020B %>% 
     filter(Source != "Total" & Sector != "Total by Fuel", NumYear == 2019) %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>% 
    select(Sector, Source, MmtCO2e)
sectormost
    
ggplot(sectormostyear, aes(NumYear,MmtCO2e, color = Sector))+
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2020 Projections By Sector")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")
    
```

```{r}
max_sector_2011 = E2011A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
    
ggplot(max_sector_2011, aes(Year,MmtCO2e, color = Sector))+
    geom_point()
    
```

#### 3.Looking at total emissions by fuel source? Which is the greatest contributor? Which has the greatest projected consumption?
```{r}
BySourceTotal = E2020B %>% 
    filter(Sector == "Total by Fuel" & Source != "Other" & Source != "Total")

BySourceTotal2019 = E2020B %>% 
    filter(Sector == "Total by Fuel" & Source != "Other" & Source != "Total" & NumYear == 2019)

ggplot(BySourceTotal2019, aes(Source,MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

ggplot(BySourceTotal, aes(NumYear, MmtCO2e, color = Source))+
    geom_line()+
    geom_smooth(se = FALSE, method = "lm")+
    ggtitle("Total Emmisions Per Fuel Source: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")
    
```
```{r}
total_source_2011 <- E2011A %>%
    filter(!str_detect(Source, "Total"))%>%
    group_by(Source)%>%
    summarize(total_emission = sum(MmtCO2e)) %>%
    arrange(desc(total_emission))
total_source_2011

source_2011 <- E2011A %>%
    filter(!str_detect(Source, "Total"))%>%
    filter(Year == "2011") %>%
    group_by(Source)%>%
    summarize(total_emission = sum(MmtCO2e)) %>%
    arrange(desc(total_emission))
source_2011

ggplot(total_source_2011, aes(x= Source, y = total_emission))+
    geom_col()
ggplot(source_2011, aes(x= Source, y = total_emission))+
    geom_col()

```


#### 4. Which sector uses the most renewables? Which uses the most non-renewables?

```{r}

C2020Renew = C2020B %>%
mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
Source != "Renewable Energy" ~ "Non-Renewable")) %>%
filter(Sector != "Energy Use & Related Statistics" & Sector != "Unspecified Sector") %>% 
group_by(Sector, RenewableResource) %>% 
drop_na() %>% 
summarise(energysum = sum(quads)) %>% 
arrange(desc(energysum))

ggplot(C2020Renew, aes(RenewableResource, energysum, fill = Sector))+
geom_col(position = "dodge")+
    ggtitle("Which Sector Uses the Most Renewables?")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Renewable")

```

```{r}

C2011Renew = C2011A %>%
    filter(Year == "2011")%>%
    filter(!str_detect(Source, "Total"))%>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
    Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector != "Energy Use & Related Statistics" & Sector != "Unspecified Sector") %>% 
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))

ggplot(C2011Renew, aes(RenewableResource, energysum, fill = Sector))+
    geom_col(position = "dodge")

```

#### 5.Which energy source is utilized the most in each sector?
```{r}
sectormostconsume = C2020B %>%
    filter(Source != "Total" & YearN == 2019) %>% 
    filter(Sector != "Energy Use & Related Statistics" & Sector != "NA") %>% 
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)

ggplot(sectormostconsume, aes(Sector, quads, fill = Source))+
geom_col()+
ggtitle("Which Fuel Source is Utilized Most in Each Sector?")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Sector")
```

```{r}
sector_most_2011 <- C2011A %>%
    filter(!str_detect(Source, "Total")) %>%
    filter(Year == "2011") %>% 
    filter(Sector != "Energy Use & Related Statistics" & Sector != "NA") %>% 
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1) 
ggplot(sector_most_2011, aes(Sector, quads, fill = Source))+
geom_col()
```

#### 6. Which sector has the highest growth rate for each fuel source?
```{r}

ratesector = C2020B %>%
    filter(YearN == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    group_by(Sector, YearN) %>% 
    slice_max(order_by = RateNumeric, n = 1) %>% 
    arrange(desc(RateNumeric)) %>% 
    select(Sector,Source,`Growth (2019-2050)`)
ratesector

```

```{r}
by_rate_2011 <- C2011A%>%
    filter(Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    drop_na()%>%
    mutate(Growth_Rate_Percentage = C2011A$`Growth Rate (2009-2035)`)%>%
    select(-`Growth Rate (2009-2035)`)%>%
    mutate(Growth_Rate_Percentage = as.numeric(sub("%","",Growth_Rate_Percentage)))%>%
    group_by(Sector,Year)%>%
    slice_max(order_by = "Growth Rate (2009-2035)")%>%
    arrange(desc(Growth_Rate_Percentage))
ggplot(by_rate_2011, aes(Sector, Growth_Rate_Percentage, fill = Source))+
    geom_col(position = "dodge")
```

#### 7. Which energy source has the highest growth rate over all? 
```{r}
ratesource = C2020B %>%
    filter(YearN == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = RateNumeric, n = 1) %>%
    select(Source,Sector,`Growth (2019-2050)`)
ratesource
```
```{r}
most_growth_2011 = by_rate_2011 %>%
    filter(Year == "2011" & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    select(Source,Sector,Growth_Rate_Percentage)%>%
    arrange(desc(Growth_Rate_Percentage))
most_growth_2011

```

#### 8. How does energy consumption vary by year?
```{r}
ConsumptionByYear = C2020B %>% 
    filter(Source =="Total") %>% 
    group_by(YearN) %>% 
    summarise(sumquads = sum(quads))

ggplot(ConsumptionByYear, aes(YearN, sumquads))+
    geom_line(color  = "red", size = 2)+
    ggtitle("Which Total Fuel Consumtion Vary by Year?")+
    ylab("Energy Consumption(quads)")+
    xlab("Year")
```
```{r}
ConsumptionByYear_2011 = by_rate %>% 
    filter(!str_detect(Source, "Total")) %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
ggplot(ConsumptionByYear_2011, aes(as.numeric(Year), sumquads))+
    geom_line()
```

#### 9.Compare projections in 2010 to actual measures in 2019. Compare slope and intercept of the trend lines.
```{r}
Actual2019 = C2020B %>% 
    filter(Year == "2019" & Source == "Total") 
ggplot(Actual2019, aes(Sector, quads, fill = Sector, alpha = 0.5))+
    geom_col()
```


### Discussion

#### Provide broader interpretations of your analysis and describe how to interpret your results with respect to your questions of interest.
#### Discuss any potential short-comings of the analysis.
#### Discuss potential future directions for additional work
#### New questions
#### Different methods to address the same questions
#### New data you might collect to refine your understanding
#### Summarize your primary conclusions and the primary evidence that supports these conclusions.
    
    
#### References
        In the R Markdown file, you may automatically cite other resources by using a caret followed by the reference between square brackets: ^[your reference text] which will place a marker at this location and a footnote at the end of the document.
