---
title: "Final Project Draft"
author: "Lindsey Fisher"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(cowplot)
library(patchwork)
library(forcats)
```

## An Analysis of US Enenrgy Related Consumption and Emmisions Data
### Group Members: Lindsey Fisher(2020 Data), Param Bhandare(2015 Data), Yunqing Xiao(2011 Data)

```{r, include=FALSE}
E2011 =read_csv("emissions2011.csv")
C2011 = read_csv("consumption2011.csv")
E2015 = read_csv("emissions2015.csv")
C2015 = read_csv("consumption2015.csv")
E2020 = read_csv("emissions2020.csv")
C2020 = read_csv("consumption2020.csv")


E2011 = E2011 %>% 
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE) 
    
E2011A = E2011 %>%
mutate(Year = as.numeric(as.character(unlist(E2011$Year))),
       GrowthRate = as.numeric(sub("%"," ",E2011$`Growth Rate (2009-2035)`))) %>% 
select(Year, Sector, Source, MmtCO2e,GrowthRate) %>% 
    filter(Sector != "NA")

C2011 = C2011 %>%
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2011 = C2011 %>% 
mutate(Year = as.numeric(as.character(unlist(C2011$Year))),
       GrowthRate = as.numeric(sub("%"," ",C2011$`Growth Rate (2009-2035)`))) %>% 
select(Year, Sector, Source, quads,GrowthRate) 


E2015 = E2015 %>% 
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "MmtCO2e",
        values_drop_na = TRUE
    )
E2015A = E2015 %>%    
    mutate(YearNum = as.numeric(as.character(unlist(E2015$Year)))) %>%
    filter(Sector != "NA")
    
C2015A = C2015 %>%
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "quads",
        values_drop_na = TRUE
    )
    
C2015B = C2015A %>% 
    mutate(Year = as.numeric(as.character(unlist(C2015A$Year))),
       GrowthRate = as.numeric(sub("%","",C2015A$`Growth (2013-2040)`))) %>% 
         select(Year, Sector,Source, quads, GrowthRate)

E2020A = E2020 %>% 
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE)
    
E2020B = E2020A %>%    
mutate(NumYear = as.numeric(as.character(unlist(E2020A$Year)))) %>% 
filter(Sector != "NA")

    

C2020A = C2020 %>%
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2020B = C2020A %>% 
mutate(Year = as.numeric(as.character(unlist(C2020A$Year))),
       GrowthRate = as.numeric(sub("%","",C2020A$`Growth (2019-2050)`))) %>% 
    select(Year, Sector, Source, quads,GrowthRate) 

```


### Introduction

We were interested in looking at something environmentally related, so we chose US energy consumption and emissions. We chose two reports from the Annual Energy Outlook (AEO) report from the US Energy Information Administration. We chose to look at the AEO published in 2011, and the AEO published in 2020. We chose these reports because they are a decade apart and will serve as a comparison to see if we have reduced emissions and consumption overall. We believe there is a lot of relationships between variables to be examined and uncovered. We are interested in relationships between the variables in each of the data sets as well as comparing the variables across the separate data sets to look for trends. We hope to use this analysis to locate key areas where the US is doing well in reducing emissions and consumption and areas where we need to focus on improvement. We hope to assess the validity of the AEOs projections by comparing the future projections in the 2011 report to the observed measures from the 2020 report. 

### Background

Emissions Data: 
The emissions data measures outputs into the atmosphere in Million Metric Ton CO2 Equivalents. This is a measure of global warming potential. CO2 is a well know gas that traps heat in the atmosphere, however, there are other gasses which trap heat at a much higher magnitude. The measurement is called CO2 equivalents because it measures the amount of total global warming potential as a factor of the magnitude of one CO2 molecules warming potential. It is separated out by sector: residential, commercial, industrial, transportation, and electric power. As well as by fuel source: petroleum, natural gas, coal, and electric power generation, and an “other” category.  The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 

Consumption Data: 
The consumption data measures how much energy was used in units of Quads. A Quad is a unit of energy equal to 10¹⁵ British Thermal Units, or 1.055 × 10¹⁸ joules in SI units. The US Department of Energy adopted the use of quads as their unit of energy use measurement for national and international reports.  The consumption data reports energy consumption by sector and power source. Sectors include residential, commercial, industrial, transportation, electric power, and delivered energy. The energy sources for consumption are more complex and vary with in each sector. The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 
    
### Analysis


#### 1. How does total emissions vary by each year? Which source is the greatest contributer for each year?

```{r}

total_2011 = E2011A %>% 
    filter(Source == "Total" & Sector == "TotalbyFuel") %>% 
    arrange(Year)

sourcetopn_2011 = E2011A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    group_by(Year) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2015 = E2015A %>% 
    filter(Source == "Total") %>% 
    arrange(YearNum)

sourcetopn2015 = E2015A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    group_by(YearNum) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2020 = E2020B %>% 
filter(Source == "Total") %>% 
arrange(NumYear)

sourcetopn2020 = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(NumYear) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

```

##### Total Emmisions By Year 

```{r}
etotal2011 = ggplot(total_2011,aes(Year,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2011 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5777, color = "red")

etotal2015 = ggplot(totalE2015, aes(YearNum,MmtCO2e)) +
    stat_smooth(se= FALSE) +
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2015 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5499.3, color = "red")

etotal2020 = ggplot(totalE2020,aes(NumYear,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("Total Emmisions Per Year: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 4993, color = "red")

etotal2011+etotal2015+etotal2020

```

##### Which source is the greatest contributer per year? How many million metric tons of CO2e are output by the greatest contributer only?

```{r}

etop2011 = ggplot(sourcetopn_2011, aes(Year, MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Fuel Source that is the Greatest Emmision Contributr By Year: AEO 2011 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(0,2500)+
geom_vline(xintercept = 2020,  color = "black", size = 2)    

etop2015 = ggplot(sourcetopn2015, aes(YearNum, MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Fuel Source that is the Greatest Emmision Contributr By Year: AEO 2015 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(0,2500)+
    geom_vline(xintercept = 2020,  color = "black", size = 2)
    
    

etop2020 = ggplot(sourcetopn2020, aes(NumYear, MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Fuel Source that is the Greatest Emmision Contributr By Year: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(0,2500)+
    geom_vline(xintercept = 2020,  color = "black", size = 2)

plot_grid(etop2011,etop2015,etop2020, nrow = 3)

```

#### 2.Which sector produces the most emissions? Does this vary by year?

```{r}
sectormost2011A = E2011A %>% 
 filter(Source != "Total" & Sector != "TotalbyFuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
    
sectormost2015A = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>%
    filter(Sector != "Carbon Dioxide Emissions") %>%
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
sectormost2020A = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
```

##### Which Sector Produces the Most Emissions?
```{r}
sectormost2011B = E2011A %>% 
     filter(Source != "Total" & Sector != "TotalbyFuel", Year == 2009) %>% 
    group_by(Sector) %>% 
    slice_max(order_by = MmtCO2e, n = 1) %>% 
    arrange(desc(MmtCO2e)) %>% 
select(Sector, MmtCO2e)
sectormost2011B

sectormost2015B = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel", Year == 2013 & Sector != "Carbon Dioxide Emissions") %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector, MmtCO2e) 
sectormost2015B

sectormost2020B = E2020B %>% 
     filter(Source != "Total" & Sector != "Total by Fuel", Year == 2019) %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector,MmtCO2e)  
sectormost2020B   

SM2011B = ggplot(sectormost2011B, aes(x = Sector,y = MmtCO2e,fill = Sector))+
    geom_col()+
    ylim(c(0,2200))
SM2015B = ggplot(sectormost2015B, aes(x = Sector,y = MmtCO2e,fill = Sector))+
    geom_col()+
    ylim(c(0,2200))
SM2020B = ggplot(sectormost2020B, aes(x = Sector,y = MmtCO2e,fill = Sector))+
    geom_col()+
    ylim(c(0,2200))

SM2011B+SM2015B+SM2020B
```

##### Variance By Year
```{r}

ggplot(sectormost2011A, aes(Year,MmtCO2e, color = Sector))+
    geom_point()+
    ggtitle("Total Emmisions Per Year By Sector: AEO 2011 ")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    geom_hline(yintercept = 1500, linetype = "dashed", color = "red")+
    scale_x_discrete(breaks = c(2010,2020,2030,2040,2050))+
    ylim(c(500,3000))


ggplot(sectormost2015A, aes(Year,MmtCO2e, color = Sector))+
geom_point()+
ggtitle("Total Emmisions Per Year By Sector: AEO 2015")+
ylab("Emissions(MmtCO2e)")+
xlab("Year")+
geom_hline(yintercept = 1500, linetype = "dashed", color = "red")+
scale_x_discrete(breaks = c(2010,2020,2030,2040,2050))+
    ylim(c(500,3000))
    
ggplot(sectormost2020A, aes(Year,MmtCO2e, color = Sector))+
    geom_point()+
    ggtitle("Total Emmisions Per Year By Sector: AEO 2020 ")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
    geom_hline(yintercept = 1500, linetype = "dashed", color = "red")+
    scale_x_discrete(breaks = c(2010,2020,2030,2040,2050))+
    ylim(c(500,3000))


```




#### 3.Looking at total emissions by fuel source, which is the greatest contributor? Which fuel source has the greatest projected consumption?

```{r}
total_source_2011 <- E2011A %>%
    filter(!str_detect(Source, "Total"))%>%
    group_by(Source)%>%
    summarize(total_emission = sum(MmtCO2e)) %>%
    arrange(desc(total_emission))

BySourceTotal2010 <- E2011A %>%
    filter(Sector == "TotalbyFuel" & Source != "Total" & Year == 2010)

 

BySourceTotal2015 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total")

BySourceTotal2014 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total" & Year == 2014)

BySourceTotal2020 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Source != "Other")

BySourceTotal2019 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Year == 2019)

```

##### Which fuel source is the greatest contribuer to emmisions?
```{r}

fuel2010 = ggplot(BySourceTotal2010, aes(x= Source, y = MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2010")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")
fuel2014 = ggplot(BySourceTotal2014, aes(Source,MmtCO2e, fill = Source))+
    geom_col() +
    ggtitle("Total Emmisions By Fuel Source: 2014")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

fuel2019 = ggplot(BySourceTotal2019, aes(Source,MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

fuel2010+fuel2014+fuel2019
```

##### Which fuel source has the greatest projected consumption?
```{r}

ggplot(BySourceTotal2020, aes(NumYear, MmtCO2e, color = Source))+
    geom_smooth(se = FALSE, method = "lm")+
    ggtitle("Total Emmisions Per Fuel Source: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")
```

#### 4. Which sector uses the most renewables? Which uses the most non-renewables?



##### 2011 DATA

```{r}

C2011Renew = C2011 %>%
    filter(Year == "2011")%>%
    filter(!str_detect(Source, "Total"))%>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
    Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector != "Energy Use & Related Statistics" & Sector != "Unspecified Sector") %>% 
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
ggplot(C2011Renew, aes(RenewableResource, energysum, fill = Sector))+
    geom_col(position = "dodge")
```

##### 2015 DATA

```{r}
C2015Renew = C2015A %>%
    slice(1:3770) %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
    Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector != "Energy Use & Related Statistics" & Sector != "Unspecified Sector") %>% 
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
ggplot(C2015Renew, aes(RenewableResource, energysum, fill = Sector))+
    geom_col(position = "dodge") +
    ggtitle("Which Sector Uses the Most Renewables?",
            subtitle = "Based on 2015 Data")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Renewable")
```

##### 2020 DATA

```{r}

C2020Renew = C2020B %>%
mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
Source != "Renewable Energy" ~ "Non-Renewable")) %>%
filter(Sector != "Energy Use & Related Statistics" & Sector != "Unspecified Sector") %>% 
group_by(Sector, RenewableResource) %>% 
drop_na() %>% 
summarise(energysum = sum(quads)) %>% 
arrange(desc(energysum))

ggplot(C2020Renew, aes(RenewableResource, energysum, fill = Sector))+
geom_col(position = "dodge")+
    ggtitle("Which Sector Uses the Most Renewables?")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Renewable")

```

#### 5.Which energy source is utilized the most in each sector?
##### 2011 DATA

```{r}
sector_most_2011 <- C2011 %>%
    filter(!str_detect(Source, "Total")) %>%
    filter(Year == "2011") %>% 
    filter(Sector != "Energy Use & Related Statistics" & Sector != "NA") %>% 
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1) 
ggplot(sector_most_2011, aes(Sector, quads, fill = Source))+
geom_col()
```

##### 2015 DATA

```{r}
sectormostconsume2015 = C2015B %>%
    slice(1:3770) %>%
    filter(Source != "Total" & Year == 2015) %>% 
    filter(Sector != "Energy Use & Related Statistics" & Sector != "NA") %>% 
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)
ggplot(sectormostconsume2015, aes(Sector, quads, fill = Source))+
    geom_col() +
    ggtitle("Which Fuel Source is Utilized Most in Each Sector?",
            subtitle = "Based on 2015 Data")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Sector")
```

##### 2020 DATA

```{r}
sectormostconsume = C2020B %>%
    filter(Source != "Total" & Year == 2019) %>% 
    filter(Sector != "Energy Use & Related Statistics" & Sector != "NA") %>% 
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)

ggplot(sectormostconsume, aes(Sector, quads, fill = Source))+
geom_col()+
ggtitle("Which Fuel Source is Utilized Most in Each Sector?")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Sector")
```

#### 6. Which sector has the highest growth rate for each fuel source?

##### 2011 DATA

```{r}
by_rate_2011 <- C2011%>%
    filter(Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    drop_na()%>%
    mutate(Growth_Rate_Percentage = C2011A$`Growth Rate (2009-2035)`)%>%
    select(-`Growth Rate (2009-2035)`)%>%
    mutate(Growth_Rate_Percentage = as.numeric(sub("%","",Growth_Rate_Percentage)))%>%
    group_by(Sector,Year)%>%
    slice_max(order_by = "Growth Rate (2009-2035)")%>%
    arrange(desc(Growth_Rate_Percentage))
ggplot(by_rate_2011, aes(Sector, Growth_Rate_Percentage, fill = Source))+
    geom_col(position = "dodge")
```

##### 2015 DATA

```{r}
ratesector2015 = C2015B %>%
    slice(1:3770) %>%
    filter(NumericYear == 2014 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    group_by(Sector, NumericYear) %>% 
    slice_max(order_by = RateNumeric, n = 1) %>% 
    arrange(desc(RateNumeric)) %>%
    select(Sector, Source, 'Growth (2013-2040)')
ratesector2015
```

##### 2020 DATA

```{r}

ratesector = C2020B %>%
    filter(YearN == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    group_by(Sector, YearN) %>% 
    slice_max(order_by = RateNumeric, n = 1) %>% 
    arrange(desc(RateNumeric)) %>% 
    select(Sector,Source,`Growth (2019-2050)`)
ratesector

```

#### 7. Which energy source has the highest growth rate over all? 

##### 2011 DATA

```{r}
most_growth_2011 = by_rate_2011 %>%
    filter(Year == "2011" & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    select(Source,Sector,Growth_Rate_Percentage)%>%
    arrange(desc(Growth_Rate_Percentage))
most_growth_2011
```

##### 2015 DATA

```{r}
ratesource = C2015B %>%
    filter(NumericYear == 2014 & Sector != "NA" & 
               Sector != "Unspecified Sector" & 
               Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = RateNumeric, n = 1) %>%
    select(Source,Sector,"Growth (2013-2040)")
```

##### 2020 DATA

```{r}
ratesource = C2020B %>%
    filter(YearN == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = RateNumeric, n = 1) %>%
    select(Source,Sector,`Growth (2019-2050)`)
ratesource
```

#### 8. How does energy consumption vary by year?

##### 2015 DATA

```{r}
ConsumptionByYear2015 = C2015B %>% 
    filter(Source =="Total") %>% 
    group_by(NumericYear) %>% 
    summarise(sumquads = sum(quads))

ggplot(ConsumptionByYear2015, aes(NumericYear, sumquads))+
    geom_line() +
    ggtitle("Which Total Fuel Consumtion Vary by Year?", 
            subtitle = "Based on 2015 Data") +
    ylab("Energy Consumption(quads)") +
    xlab("Year")
```

##### 2011 DATA

```{r}
ConsumptionByYear_2011 = by_rate_2011 %>% 
    filter(!str_detect(Source, "Total")) %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
ggplot(ConsumptionByYear_2011, aes(as.numeric(Year), sumquads))+
    geom_line()
```

##### 2020 DATA

```{r}
ConsumptionByYear2020 = C2020B %>% 
    filter(Source =="Total") %>% 
    group_by(YearN) %>% 
    summarise(sumquads = sum(quads))

ggplot(ConsumptionByYear2020, aes(YearN, sumquads))+
    geom_line(color  = "red", size = 2)+
    ggtitle("Which Total Fuel Consumtion Vary by Year?")+
    ylab("Energy Consumption(quads)")+
    xlab("Year")
```

#### 9.Compare projections in 2010 to actual measures in 2019. Compare slope and intercept of the trend lines.

```{r}
Actual2019 = C2020B %>% 
    filter(Year == "2019" & Source == "Total" & Sector != "Unspecified Sector") 

Prediction2015 = C2015B %>% 
    filter(Year == "2019" & Source == "Total" & Sector != "Unspecified Sector" & Sector != "Total Energy Consumption")

Prediction2011 = C2011 %>% 
    filter(Year == "2019" & Source == "Total" & Sector != "Total Energy Consumption")

ggplot(Actual2019, aes(Sector, quads, fill = Sector))+
    geom_col()

ggplot(Prediction2011, aes(Sector, quads, fill = Sector))+
geom_col()
    
ggplot(Prediction2015, aes(Sector, quads, fill = Sector))+    
geom_col()

```

```{r}
ConsumptionByYear2015 <- ConsumptionByYear2015 %>%
    mutate(Dataset = "2015")
ConsumptionByYear2020 <- ConsumptionByYear2020 %>%
    mutate(Dataset = "2020")
consumptions_2015_2020_combined <- bind_rows(ConsumptionByYear2015, ConsumptionByYear2020)
sourcetopn2015 <- sourcetopn2015 %>%
    mutate(Dataset = "2015")
sourcetopn2020 <- sourcetopn2020 %>%
    mutate(Dataset = "2020")
emission_2015_2020_combined <- bind_rows(sourcetopn2015, sourcetopn2020)

ggplot(consumptions_2015_2020_combined, aes(NumericYear, sumquads, color = Dataset)) +
geom_point()
    
ggplot(emission_2015_2020_combined, aes(YearNum, MmtCO2e, color = Dataset)) +
    geom_point()
```

### Discussion

Param:
Broadly speaking, based on the 2015 and 2020 datasets, it is certain that emissions have dropped. However, it is also evident that consumption will either mostly stay steady or rise slowly. Projections for consumption skyrocket in each dataset regardless of year, so it does appear the AEO is taking the pessimistic outlook for their projections. 

The datasets for 2015 energy outlook seems to be slightly inconsistent with format and depth of projections with other datasets. This is not too much of a shortcoming, but as analysts, it does leave us guessing for the verification of some of the prior projections in later datasets within a single dataset. Additionally, the outlook is made during the fiscal year that the outlook is released, so only the previous few years will be accurate. This requires more data to be obtained, which in turn requires more standardization between the format of each dataset. Perhaps, given the previous years data-sets and actual values, a match between the mesh of datas that are equal could come handy. And to test performance even further, a whole parity chart could be made between different datasets actual performance and projected performance.

Further analysis can be made on specific governmental body's policy performance, specifically around the time period they are made in tandem to the associated energy outlook data sets. Similarly, new data can be collected regarding this.

Conclusively, it appears Americans are doing a very good job on curbing emissions when considering the 2015 and 2020 dataset. Consumption is still there, but it is highly unlikely an increasing population will decrease their consumption of energy unless several billions of technology's source of energy is changed to a more modern source, this also requiring a change in the technology itself.

Yunqing:
- for data in 2011 delivered energy consumptions contribute the most among all sectors of energy and coal is the most commonly used source of energy followed by electricity and petroleium. Transportation and electric power produces the most parts of the energy.And a huge portion of the energy is used for US GDP. Through these comparision, it will be interesting to look at how it changed trhoughout the decade.
- one of the potential short-comings of this analysis will be the sized of our datasets are slightly different which might caused a little inaccuracy in final analysis
- we can look more deeply into how different variables changes throughout the decades and discuss the potential caused and effect of the changes such as how the a environmental related policy or crisis impacted the energy use or how some major changes in economy or manufacturing market affect the use of energy
 - my new questions is can we use some of the statistc model to predict the future changes based on previous year's dataset which we can later use to compare with the actual situation and talk about why there are potential difference between prediction and the reality
 - Instead of comparing all sectors and sources, I will like to look closely at a specific source and sector and study it more in-depth
- I will like to look at some data that related to the variable which changed the most throughout the decade and study the reason behind the change
- In conclusion, based on the previous analysis, I would like to predict that the consumption of energy continue to increase in the next decade but other source of the energy will surpase the domination of energy use of coal becasue we can see the curve of electric power increasing with a higher slope the coal in 2011