---
title: "An Analysis of US Energy Related Consumption and Emissions Data"
author: "Lindsey Fisher,Param Bhandare,Yunqing Xiao"
date: "12/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(patchwork)
library(forcats)
source("viridis.R")
source("ggprob.R")
library(gridExtra)
```


```{r, include=FALSE}
E2011 =read_csv("emissions2011.csv")
C2011 = read_csv("consumption2011.csv")
E2015 = read_csv("emissions2015.csv")
C2015 = read_csv("consumption2015.csv")
E2020 = read_csv("emissions2020.csv")
C2020 = read_csv("consumption2020.csv")

E2011 = E2011 %>% 
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE) 
    
E2011A = E2011 %>%
mutate(Year = as.numeric(as.character(unlist(E2011$Year))),
       GrowthRate = as.numeric(sub("%"," ",E2011$`Growth Rate (2009-2035)`))) %>% 
select(Year, Sector, Source, MmtCO2e,GrowthRate) %>% 
    filter(Sector != "NA")

C2011 = C2011 %>%
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2011 = C2011 %>% 
mutate(Year = as.numeric(as.character(unlist(C2011$Year))),
       GrowthRate = as.numeric(sub("%"," ",C2011$`Growth Rate (2009-2035)`))) %>% 
  filter(Sector != "Energy Use and Related Statistics	") %>% 
select(Year, Sector, Source, quads,GrowthRate) 


E2015 = E2015 %>% 
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "MmtCO2e",
        values_drop_na = TRUE
    )
E2015A = E2015 %>%    
    mutate(Year = as.numeric(as.character(unlist(E2015$Year)))) %>%
    filter(Sector != "NA")
    
C2015A = C2015 %>%
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "quads",
        values_drop_na = TRUE
    )
    
C2015B = C2015A %>% 
    mutate(Year = as.numeric(as.character(unlist(C2015A$Year))),
       GrowthRate = as.numeric(sub("%","",C2015A$`Growth (2013-2040)`))) %>% 
         select(Year, Sector,Source, quads, GrowthRate)

E2020A = E2020 %>% 
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE)
    
E2020B = E2020A %>%    
mutate(Year = as.numeric(as.character(unlist(E2020A$Year)))) %>% 
filter(Sector != "NA")

C2020A = C2020 %>%
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2020B = C2020A %>% 
mutate(Year = as.numeric(as.character(unlist(C2020A$Year))),
       GrowthRate = as.numeric(sub("%","",C2020A$`Growth (2019-2050)`))) %>% 
    select(Year, Sector, Source, quads,GrowthRate) 

EnergyUseStat2020 = C2020B %>% 
  filter(Sector == "Energy Use & Related Statistics" | Source == "Carbon Dioxide Emissions (million metric tons carbon dioxide)")

```


# Introduction

  Our group was interested in looking at an environmentally related issue as green energy and climate change is a hot topic for debate currently. We will use this report to analyze United States energy consumption and emissions data.We are using energy measurements provided by the Annual Energy Outlook (AEO) report which is published each year by the the US Energy Information Administration. We chose to look at  AEO reports published in 2011, 2015 and 2020 since they within one decade and we will be able to compare across these years to see im we have reduced emissions and consumption overall in the past decade.
  
  For this report we will visually explore relationships between energy data metrics while also looking at our progress as a country for reducing emissions and consumption overall. We believe there may be interesting relationships between variables as well as among variables from separate data sets to be uncovered through visual analysis and modeling. We are interested in relationships between the variables such as: Industry Sector, Fuel Source, Emissions, and Consumption metircs within each of the data sets as well as comparing the variables across the separate years to look for trends.
  
  We will use this analysis to locate key areas where the US is suceeding at reducing emissions and areas where we need to improvement. We will assess the validity of the AEOs projections by comparing the future projections from the 2010 and 2015 reports to the observed measures supplied by the 2020 report. 
  
  In order to accomplish our goal of identifying if the US is reducing emissions, we will compare the prediction of energy use for the year 2020 obtained from the 2010 and 2015 AEO Reports against the actual data from 2020. Also, we will look at individual energy sources and sectors to better understand how these overall changes are being influence by sector level changes.

# Background

 The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 
 
**Emissions Data:**

The emissions data measures outputs into the atmosphere in units of Million Metric Ton CO2 Equivalents. This is a measure of global warming potential. CO2 is a well know gas that traps heat in the atmosphere, however, there are other gasses which trap heat at a much higher magnitude. The measurement is called CO2 equivalents because it measures the amount of total global warming potential as a factor of the magnitude of one CO2 molecules warming potential. 

Variables: 

*Sectors* 
Includes Residential, Industrial, Transportation, Commercial, Electric Power

*Fuel Source*

*Year*

*Growth Rate*

*MmtCo2e*
Emissions measured in units of Million metric tons of Carbon Dioxide Equivalents

**Consumption Data:**

The consumption data measures how much energy was used in units of Quads. A Quad is a unit of energy equal to 10¹⁵ British Thermal Units, or 1.055 × 10¹⁸ joules in SI units. The US Department of Energy adopted the use of quads as their unit of energy use measurement for national and international reports.  The consumption data reports energy consumption by sector and power source. Sectors include residential, commercial, industrial, transportation, electric power, and delivered energy. The energy sources for consumption are more complex and vary with in each sector. 
 
Variables: 

*Sector* 

*Fuel Source*

*Year*

*Growth Rate*

*Quads*
A Quad is a unit of energy equal to 10¹⁵ British Thermal Units, or 1.055 × 10¹⁸ joules in SI units. 


We choose these data because they clearly presents the different meausres for emission and consumption of energy with labeled year, sector(the area that energy is being used) and source of energy. Thus, in order to see whether the US is doing well in reducing emissions and consumption and areas where we need to focus on improvement, we need to use these information to analyze and discuss about the energy performance in the US. 
 
 
The energy used in the consumption data are measure by MmtCO2e (Million Metric Ton CO2) which is a common unit for measuring the energy emission and consumption. The energy used in the emission data are meausred by quads. These datasets divided the energy by its sector and source which makes the anlysis easier because we can look at different energy sector and source separately to see which of them contribute the most in the increasing of emission and consumption of energy and which of them reduced the most. Also, we choose to use the data within a 10-year period because not only the common energy source and how energy is used are similar within 10 year, but also because we can track the changes in different variables closely in a shorter period and time and compare the actual changes in 10 years with our prediction from the previous dataset.


# Analysis


## **(1) How do total emissions vary by year?**
####The intersecting red lines allow us to compare total emissions for 2020 as predicted by each AEO report.
####These graphics show that actual total emissions in 2020 is much lesser than the predicted total emissions by the 2011 outlook & 2015 outlook reports. 

```{r,echo = FALSE}

total_2011 = E2011A %>% 
    filter(Source == "Total" & Sector == "TotalbyFuel") %>% 
    arrange(Year)

sourcetopn_2011 = E2011A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
  filter(Year == 2010) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2015 = E2015A %>% 
    filter(Source == "Total") %>% 
    arrange(Year)

sourcetopn2015 = E2015A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    filter(Year == 2014)%>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2020 = E2020B %>% 
filter(Source == "Total") %>% 
arrange(Year)

sourcetopn2020 = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
  filter(Year == 2019) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

```


```{r,echo = FALSE}
etotal2011 = ggplot(total_2011,aes(Year,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("2011 Data")+
    ylab(" Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5777, color = "red")

etotal2015 = ggplot(totalE2015, aes(Year,MmtCO2e)) +
    stat_smooth(se= FALSE) +
    geom_point()+
    ggtitle("2015 Data")+
    ylab(" Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5499.3, color = "red")

etotal2020 = ggplot(totalE2020,aes(Year,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("2020 Data")+
    ylab("Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 4993, color = "red")

grid.arrange(etotal2011,etotal2015,etotal2020, ncol = 3)

```

## **(2) Which fuel source is the greatest emissions contributer? How many million metric tons of CO2e are output by the greatest contributer?**
####Petroleum is the fuel source that has contributed most to emissions over all three reports. 
####When comparing 2010 to 2019 we can see a decrease in overall emissions from petrolem. 

```{r,echo = FALSE}

etop2011 = ggplot(sourcetopn_2011, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2010")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)   

etop2015 = ggplot(sourcetopn2015, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2014")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)
  
    
etop2020 = ggplot(sourcetopn2020, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)


grid.arrange(etop2011,etop2015,etop2020, ncol = 3)

```

## **(3)Which sector produces the most emissions? Does this vary by which year the outlook report is pulished?**
####The order of sectors, highest emiiting to lowest emitting sectors, is the same for the year 2010 and 2019 although there are some slight fluxuations in the ammount of emissions produced by each sector. 

```{r,echo = FALSE}
sectormost2011A = E2011A %>% 
 filter(Source != "Total" & Sector != "TotalbyFuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
    
sectormost2015A = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>%
    filter(Sector != "Carbon Dioxide Emissions") %>%
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
sectormost2020A = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
```


```{r,echo = FALSE}
sectormost2011B = E2011A %>% 
     filter(Source != "Total" & Sector != "TotalbyFuel", Year == 2009) %>% 
    group_by(Sector) %>% 
    slice_max(order_by = MmtCO2e, n = 1) %>% 
    arrange(desc(MmtCO2e)) %>% 
select(Sector, MmtCO2e) 


sectormost2015B = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel", Year == 2013 & Sector != "Carbon Dioxide Emissions") %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector, MmtCO2e) 

sectormost2020B = E2020B %>% 
     filter(Source != "Total" & Sector != "Total by Fuel", Year == 2019) %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector,MmtCO2e) 
   


SM2011B = ggplot(sectormost2011B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
  ggtitle("AEO 2011")+
  xlab("Sector")+
    ylim(c(0,2200))
SM2015B = ggplot(sectormost2015B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
   ggtitle("AEO 2015")+
  xlab("Sector")+
    ylim(c(0,2200))
SM2020B = ggplot(sectormost2020B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
   ggtitle("AEO 2020")+
  xlab("Sector")+
    ylim(c(0,2200))

grid.arrange(SM2011B, SM2015B, SM2020B,nrow = 3)

```


## **(4)Looking at total emissions by fuel source, which is the greatest contributor? Which fuel source has the greatest projected emissions?**


```{r,echo = FALSE}
total_source_2011 <- E2011A %>%
    filter(!str_detect(Source, "Total"))%>%
    group_by(Source)%>%
    summarize(total_emission = sum(MmtCO2e)) %>%
    arrange(desc(total_emission))

BySourceTotal2010 <- E2011A %>%
    filter(Sector == "TotalbyFuel" & Source != "Total" & Year == 2010)

 

BySourceTotal2015 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total")

BySourceTotal2014 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total" & Year == 2014)

BySourceTotal2020 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Source != "Other")

BySourceTotal2019 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Year == 2019)

```

### Which fuel source is the greatest contribuer to emmisions?
####Petroleum is the greatest contributer within all three reports. However, natural gas surpassed coal for the second greatest emitter in 2014 and 2019.
```{r,echo = FALSE}

fuel2010 = ggplot(BySourceTotal2010, aes(x= Source, y = MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2010")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")
fuel2014 = ggplot(BySourceTotal2014, aes(Source,MmtCO2e, fill = Source))+
    geom_col() +
    ggtitle("Total Emmisions By Fuel Source: 2014")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

fuel2019 = ggplot(BySourceTotal2019, aes(Source,MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

grid.arrange(fuel2010,fuel2014,fuel2019)


```

### Which fuel source has the greatest projected emissions?
#### Coal and Petroleum are both projected to reduce their emmisions from 2020 to 2050 while natural gas is projected to increase its emissions. 

```{r,echo = FALSE}

ggplot(BySourceTotal2020, aes(Year, MmtCO2e, color = Source))+
    geom_smooth(se = FALSE, method = "lm")+
    ggtitle("Total Emmisions Per Fuel Source: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
  theme(element_blank())
```

## **(5) Which sector uses the most renewables? Which uses the most non-renewables?**
#### For the 2020 data we can see that the industrial sector uses the most renewables whereas the transportation sector uses the most non renewables. Comparing the 2011 data to the 2020 data we can see that there has been a significant change in the renewability of resources that the industrial sector uses. 

```{r,echo = FALSE}

C2011Renew = C2011 %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable", Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>% 
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))

C2015Renew = C2015A %>%
    slice(1:3770) %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
                                        Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
C2020Renew = C2020B %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
    Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
```

```{r,echo = FALSE}
renew_2011=ggplot(C2011Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", fill = "lightblue")+
  facet_wrap(~RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
            subtitle = "Based on 2011 Data")+
     ylab("Consumptionr(quads)")+
    xlab("Sector")+
  ylim(0,3000)+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    
renew_2015=ggplot(C2015Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", fill = "blue") +
    facet_wrap(~ RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
            subtitle = "Based on 2015 Data")+
    ylab("Consumptionr(quads)")+
    xlab("Renewable")+
  ylim(0,3000)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
renew_2020 = ggplot(C2020Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", fill = "navy")+
    facet_wrap(~RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
        subtitle = "Based on 2020 Data")+
    ylab("Consumptionr(quads)")+
    xlab("Renewable")+
  ylim(0,3000)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

grid.arrange(renew_2011,renew_2015,renew_2020, ncol =3)

```

## **(6) Which source consumes the most energy in each sector?**
#### From this graph we can conclude that electricty and delivered energy play a major role in the ammount of energy we consume across all sectors. 
```{r,echo = FALSE}
sector_most_2011 <- C2011 %>%
    filter(!str_detect(Source, "Total")) %>%
    filter(Year == "2011") %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)

sectormostconsume2015 = C2015B %>%
    slice(1:3770) %>%
    filter(Source != "Total" & Year == 2015) %>% 
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)
sectormostconsume2020 = C2020B %>%
    filter(Source != "Total" & Year == 2019) %>% 
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector) %>%
    slice_max(order_by = quads, n = 1)
most_used_by_sector = bind_rows(sector_most_2011,sectormostconsume2015,sectormostconsume2020)
```

```{r,echo = FALSE}
ggplot(most_used_by_sector, aes(Sector, quads, fill = Source) ) + 
    geom_col()+
    facet_wrap( ~ Year, nrow = 1) + 
    ggtitle("Which Fuel Source is Utilized Most in Each Sector?")+
    ylab("Energy Consumption per Sector(quads)")+
    xlab("Sector")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## **(7) Which energy source has the highest growth rate over all?**
#### This graph shows how we have transitioned from coal power, to natural gas, and how we are currently growing into new methods to supply our energy such as hydrogen fuel sources. 

```{r,echo = FALSE}
ratesource2011 = C2011 %>%
    filter(Year == 2010 & Sector != "NA") %>% 
    filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" | Sector == "Comericial" | Sector == "Delivered Energy Consumption" | Sector == "Electric Power ") %>% 
    select(Source,Sector,GrowthRate,Year)%>%
    slice_max(order_by = GrowthRate, n = 1)

ratesource2015 = C2015B %>%
    filter(Year == 2014 & Sector != "NA" & 
               Sector != "Unspecified Sector" & 
               Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate,Year)

ratesource2020 = C2020B %>%
    filter(Year == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate,Year)

highgrowthrate = bind_rows(ratesource2011,ratesource2015,ratesource2020)

ggplot(highgrowthrate,aes(Year,GrowthRate, fill = Source))+
    geom_col()+
     theme(
  # Hide panel borders and remove grid lines
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  axis.line = element_line(colour = "black"))+
    annotate("text", x=c(2010,2014,2019), y=5, label = c("2010","2014","2019"), color = "black", size = 5)+
  scale_x_discrete()

```

## **(8)How do energy consumption projection trends change based on the year the report was published?**
#### From each report we can see that the energy consuption trend is positive and linear. However, it is important to focus on the y-axis, as we have lowered our projected consumption for 2030 by nearly 50 quads. 

```{r,echo = FALSE}
ConsumptionByYear_2011 = C2011 %>% 
    filter(Source == "Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
ConsumptionByYear2015 = C2015B %>% 
    filter(Source == "Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
ConsumptionByYear2020 = C2020B %>% 
    filter(Source =="Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
q82011 = ggplot(ConsumptionByYear_2011, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2011 Data") +
    ylab("quads") +
    xlab("Year")+
    xlim(2020,2050)+
    ylim(225,375)+
     theme_dark()
q82015 = ggplot(ConsumptionByYear2015, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2015 Data") +
    ylab("quads") +
    xlab("Year")+
    xlim(2020,2050)+
    ylim(225,375)+
     theme_dark()
q82020 = ggplot(ConsumptionByYear2020, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2020 Data") +
    ylab("quads") +
    xlab("Year")+
      xlim(2020,2050)+
    ylim(225,375)+
    theme_dark()

grid.arrange(q82011,q82015,q82020)
   
    
    
```

## **(9).Compare projections for 2019 agaisnt to actual measures in 2019.**

```{r,echo = FALSE}
Actual2019 = C2020B %>% 
    filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ")
Prediction2015 = C2015B %>% 
        filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ")
Prediction2011 = C2011 %>% 
       filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ")
ggplot(Actual2019, aes(x = reorder(Sector, -quads), quads))+
    geom_col(fill = "lightblue")+
  ggtitle("Actual Values For Year 2019", subtitle ="From AEO 2020")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)
  
ggplot(Prediction2011, aes(x = reorder(Sector, -quads), quads))+
geom_col(fill = "blue")+
  ggtitle("Predicted Values For Year 2019", subtitle =  "From AEO 2011")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)
    
ggplot(Prediction2015, aes(x = reorder(Sector, -quads), quads))+    
geom_col(fill = "navy")+
  ggtitle("Predicted Values For Year 2019", subtitle =  "From AEO 2015")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)
```

```{r, echo = FALSE}
ConsumptionByYear2015 <- ConsumptionByYear2015 %>%
    mutate(Dataset = "2015")
ConsumptionByYear2020 <- ConsumptionByYear2020 %>%
    mutate(Dataset = "2020")
consumptions_2015_2020_combined <- bind_rows(ConsumptionByYear2015, ConsumptionByYear2020)
sourcetopn2015 <- sourcetopn2015 %>%
    mutate(Dataset = "2015")
sourcetopn2020 <- sourcetopn2020 %>%
    mutate(Dataset = "2020")
emission_2015_2020_combined <- bind_rows(sourcetopn2015, sourcetopn2020)
ggplot(consumptions_2015_2020_combined, aes(Year, sumquads, fill = Dataset)) +
geom_col()
    
ggplot(emission_2015_2020_combined, aes(Year, MmtCO2e, fill = Dataset)) +
    geom_col()
```

## **(10) What are the projected trends for GDP, Total Energy Consumption, Emissions, and Population? Do these variable seem to interact over the next 10 years?**
####We can visualize several relationships from this figure:
####-GDP and Population increase in a positive linear direction - Emissions is projected to decrease even though popultion is increasing. This could imply that we are using fuels that emmit less per person. 
####-The trend for total energy consumption is interesting and could use further research. 
```{r}

EUS2020GDP = EnergyUseStat2020 %>% 
  filter(Source == "Gross Domestic Product (billion 2012 dollars)") %>%  
  pivot_wider(names_from = Source, values_from = quads) %>% 
  rename(GDP = "Gross Domestic Product (billion 2012 dollars)") %>% select(Year,GDP)

EUS2020CDE = EnergyUseStat2020 %>% 
filter(Source == "Carbon Dioxide Emissions (million metric tons carbon dioxide)") %>%  
pivot_wider(names_from = Source, values_from = quads) %>% 
  rename(CO2E = "Carbon Dioxide Emissions (million metric tons carbon dioxide)") %>% 
select(Year,CO2E)

EUS2020POP = EnergyUseStat2020 %>% 
  filter(Source == "Population (millions)") %>%  
  pivot_wider(names_from = Source, values_from = quads) %>% 
  rename(population = "Population (millions)") %>% 
  select(Year,population) 
EUS2020Energy = EnergyUseStat2020 %>% 
  filter(Source == "Total Energy Use") %>% 
  pivot_wider(names_from = Source, values_from = quads) %>% 
  select(Year,`Total Energy Use`) 
eustats = inner_join(EUS2020CDE,EUS2020Energy) %>% 
  inner_join(EUS2020GDP) %>% 
  inner_join(EUS2020POP) %>% 
  filter(Year == 2020 |Year == 2021|Year == 2022|Year == 2023|Year == 2024|Year == 2025|Year == 2026|Year == 2027|Year == 2028|Year == 2029|Year == 2030)


AA = ggplot(eustats, aes(Year,GDP))+
geom_point()+
  geom_smooth()+
  scale_x_discrete(breaks = c(2020,2030,2040,2050))+
  ylab("GDP(billion 2012 dollars)")

BB = ggplot(eustats, aes(Year,population))+
geom_point()+
  geom_smooth()+
  scale_x_discrete(breaks = c(2020,2030,2040,2050))+
  ylab("Population(Millions")
CC = ggplot(eustats, aes(Year,CO2E))+
geom_point()+
  geom_smooth()+
  scale_x_discrete(breaks = c(2020,2030,2040,2050))+
  ylab("million metric tons CO2")
DD = ggplot(eustats, aes(Year,`Total Energy Use`))+
geom_point()+
  geom_smooth()+
  scale_x_discrete(breaks = c(2020,2030,2040,2050))+
  ylab("Total Energy Consumption (Quads)")

grid.arrange(AA,BB,CC,DD, nrow= 2, ncol=2)

```

## **(12)How does the use of Petroleum and Coal changed from 2010-2019 period to 2020-2030 period?**

### Petroleum

 - From the previous table, we can see that Petroleum is a main source of energy contributing to the energy emission. As the technology and use of energy changed and developed throughout the decase, we are interested in how the use of petroleum changes in the sector of transportation.  From the previous problem, we can see that the main sector uses petroleum is in transportation. We now want to compare the result within 20 dacades to see whether the mean use of Coal for emission changes significantly.

```{r echo=FALSE}
p_c2010 = E2011A%>%
    filter(Source == "Petroleum" & Sector != "Total by Fuel")%>%
    filter(Year< 2020)
p_c2020 = E2020B%>%
    filter(Source == "Petroleum" & Sector != "Total by Fuel")%>%
    filter(Year < 2030 & Year >= 2020)
p_c = bind_rows(p_c2010,p_c2020)%>%
    filter(Sector == "Transportation")%>%
    mutate(range = case_when(Year < 2019 ~ "2010",
                     Year >= 2019 ~ "2020"))
sum_PC = p_c%>%
    group_by(range)%>%
    summarize(n = n(),
              sum_MmtCO2e = sum(MmtCO2e),
              mean = mean(sum_MmtCO2e),
              sd = sd(MmtCO2e))
ggplot(p_c, aes(x=log10(MmtCO2e),color=range)) +
  geom_density() +
  xlab("MmtCO2e(log10)") +
  ylab("Density") +
  ggtitle("Mean source of energy in Emission (MmtCO2e)",
          subtitle="early = 2010-2020, recent = 2020-2030") +
  theme_bw()
```
- Thus, We are 95% confident that the mean emission energy from Petroleum meausred by MmtCO2e in period 2009-2018 was between 227.9347 MmtCO2e lower to 296.1169 MmtCO2e higher than the mean emission energy from Petroleum meausred by MmtCO2e in period 2019-2029.

 - Now, using the mean and se generated from the previous t-test, we now are able to verify the p value by calculating the test statstistics.
```{r echo=FALSE}
tstat_p <- mean_diff_p/se_p
## pvalue
pt(-abs(tstat_p), dof_p) + pt(abs(tstat_p), dof_p, lower.tail=FALSE)
```
 - The p value is 0.0002402814 (less than 0.05 (typically ≤ 0.05)) is statistically significant. It indicates strong evidence against the null hypothesis, as there is less than a 5% probability the null is correct (and the results are random). Therefore, we reject the null hypothesis, and accept the alternative hypothesis of the mean use of petroleum in transportation is different from 2010-2020 period to 2020-2030 period. 
### Coal

- From the previous table, we can see that Coal is also a main source of energy contributing to the energy emission. Also, we noticed that the major use of Coal is in Sector of Electric Power. As the technology and use of energy changed and developed throughout the decase, we are instrerested in how the use of Coal changes in the sector of electric power. We now want to compare the result within 20 dacades to see whether the mean use of Coal for emission in Electric Power changes significantly.
```{r echo=FALSE}
c2010 = E2010A%>%
    filter(Source == "Coal" & Sector != "Total by Fuel")%>%
    filter(Sector == "Electric Power")%>%
    filter(Year < 2019)
c2020 = E2020B%>%
    filter(Source == "Coal" & Sector != "Total by Fuel")%>%
    filter(Sector == "Electric Power")%>%
    filter(Year < 2030 & Year >=2020)
c = bind_rows(c2010,c2020)%>%
    mutate(range = case_when(Year < 2019 ~ "2010",
                     Year >= 2019 ~ "2020"))
sum_C = c%>%
    group_by(range)%>%
    summarize(n = n(),
              sum_MmtCO2e = sum(MmtCO2e),
              mean = mean(sum_MmtCO2e),
              sd = sd(MmtCO2e))
ggplot(c, aes(x=log10(MmtCO2e),color=range)) +
  geom_density() +
  xlab("MmtCO2e(log10)") +
  ylab("") +
  ggtitle("Coal Use in Electric Power for Energy Emission",
          subtitle="early = 2010-2020, recent = 2020-2030") +
  theme_bw()
```
#### Test the hypothesis that the mean energy use from Coal in 2010-2020 (Pa) is equal to that in 2020-2030 period (Pb) versus that alternative that they are different. 

$$
H_0: P_a = P_b
$$

$$
H_a:  P_a \neq P_b
$$
```{r echo=FALSE}
x1_c <- c %>%
  filter(range == "2010") %>%
  pull(MmtCO2e)
x2_c <- c %>%
  filter(range == "2020") %>%
  pull(MmtCO2e)
```
```{r echo=FALSE}
t.test(x1_c,x2_c)
mean_diff_c <- mean(x1_c) - mean(x2_c)
dof_c <- 16.803 ## from t.test() above
se_c <- sqrt(sd(x1_c)^2/length(x1_c) + sd(x2_c)^2/length(x2_c))
ci_5_c <- mean_diff_c + c(-1,1)*qt(0.975,dof_c)*se_c
signif(ci_5_c,3)
```
We are 95% confident that the mean emission energy from Coal meausred by MmtCO2e in period 2009-2018 was between 37.6 MmtCO2e lower to 501.0 MmtCO2e higher than the mean emission energy from Coal meausred by MmtCO2e in period 2019-2029.


Test the hypothesis that the mean energy use from Coal in 2010-2019 period is equal to that in the recent era versus that alternative that they are different. 

```{r echo=FALSE}
tstat_c <- mean_diff_c/se_c
## pvalue
pt(-abs(tstat_c), dof_c) + pt(abs(tstat_c), dof_c, lower.tail=FALSE)
```
 - The p value is 4.151938e-51 (less than 0.05 (typically ≤ 0.05)) is statistically significant. It indicates strong evidence against the null hypothesis, as there is less than a 5% probability the null is correct (and the results are random). Therefore, we reject the null hypothesis, and accept the alternative hypothesis of the mean use of Coal in electric power is different from 2010-2020 period to 2020-2030 period. As renewable and other more efficient energy source being used, from the graph in previous question, we can seethe use of coal decline in electric power which contributing much lesser in energy emission. This is a good news because the coal-fired power plants release more greenhouse gases per unit of energy produced than any other electricity source (according to https://www.greenamerica.org/fight-dirty-energy/amazon-build-cleaner-cloud/coal-why-it-dirty).

#### Test the hypothesis that the mean energy use from Petroleum in 2010-2020 (Pa) is equal to that in 2020-2030 period (Pb) versus that alternative that they are different. 

$$
H_0: P_a = P_b
$$

$$
H_a:  P_a \neq P_b
$$
```{r echo=FALSE}
x1_p <- p_c %>%
  filter(range == "2010") %>%
  pull(MmtCO2e)
x2_p <- p_c %>%
  filter(range == "2020") %>%
  pull(MmtCO2e)
t.test(x1_p,x2_p)
mean_diff_p <- mean(x1_p) - mean(x2_p)
dof_p <- 15.664 ## from t.test() above
se_p <- sqrt(sd(x1_p)^2/length(x1_p) + sd(x2_p)^2/length(x2_p))
ci_5_p <- mean_diff_p + c(-1,1)*qt(0.975,dof_p)*se_p
```

## Discussion
