---
title: "Final Project Draft"
author: "Lindsey Fisher"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(cowplot)
library(patchwork)
library(forcats)
source("viridis.R")
source("ggprob.R")
library(gridExtra)
```

## An Analysis of US Enenrgy Related Consumption and Emmisions Data
##### Group Members: Lindsey Fisher(2020 Data), Param Bhandare(2015 Data), Yunqing Xiao(2011 Data)

```{r, include=FALSE}
E2010 = read_csv("emissions2010.csv")
C2010 = read_csv("consumption2010.csv")
E2015 = read_csv("emissions2015.csv")
C2015 = read_csv("consumption2015.csv")
E2020 = read_csv("emissions2020.csv")
C2020 = read_csv("consumption2020.csv")

E2010 = E2010 %>%
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE)

E2010A = E2010 %>%
mutate(Year = as.numeric(as.character(unlist(E2010$Year))),
       GrowthRate = as.numeric(sub("%"," ",E2010$`Growth (2008-2035)`))) %>%
select(Year, Sector, Source, MmtCO2e,GrowthRate) %>%
    filter(Sector != "NA")
C2010A = C2010 %>%
    pivot_longer(
   cols = c(3:30),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)
C2010B = C2010A %>%
mutate(Year = as.numeric(as.character(unlist(C2010A$Year))),
       GrowthRate = as.numeric(sub("%","",C2010A$X32))) %>%
    select(Year, Sector, Source, quads,GrowthRate)


E2015 = E2015 %>% 
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "MmtCO2e",
        values_drop_na = TRUE
    )
E2015A = E2015 %>%    
    mutate(YearNum = as.numeric(as.character(unlist(E2015$Year)))) %>%
    filter(Sector != "NA")
    
C2015A = C2015 %>%
    pivot_longer(
        cols = c(3:31),
        names_to = "Year",
        names_prefix = "Year",
        values_to = "quads",
        values_drop_na = TRUE
    )
    
C2015B = C2015A %>% 
    mutate(Year = as.numeric(as.character(unlist(C2015A$Year))),
       GrowthRate = as.numeric(sub("%","",C2015A$`Growth (2013-2040)`))) %>% 
         select(Year, Sector,Source, quads, GrowthRate)

E2020A = E2020 %>% 
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "MmtCO2e",
   values_drop_na = TRUE)
    
E2020B = E2020A %>%    
mutate(NumYear = as.numeric(as.character(unlist(E2020A$Year)))) %>% 
filter(Sector != "NA")

    

C2020A = C2020 %>%
    pivot_longer(
   cols = c(3:34),
   names_to = "Year",
   names_prefix = "Year",
   values_to = "quads",
   values_drop_na = TRUE)

C2020B = C2020A %>% 
mutate(Year = as.numeric(as.character(unlist(C2020A$Year))),
       GrowthRate = as.numeric(sub("%","",C2020A$`Growth (2019-2050)`))) %>% 
    select(Year, Sector, Source, quads,GrowthRate) 

```


### Introduction

 - For the past decade, the use of energy became more and more important in our daily life, and the modes and source of energy using and mission has changed throughout the year. Thus, We were interested in looking at something environmentally related, so we chose US energy consumption and emissions. We chose two reports from the Annual Energy Outlook (AEO) report from the US Energy Information Administration. We chose to look at the AEO published in 2010, 2015, and 2020 because they within one decade and will serve as a comparison to see if we have reduced emissions and consumption overall in the past decade. 
 - We believe there is a lot of relationships between variables to be examined and uncovered. We are interested in relationships between the variables in each of the data sets as well as comparing the variables across the separate data sets to look for trends. For example, how does the major source of energy changes within the decade, and how does the total use of energy changes in the past 10 year.
 - We hope to use this analysis to locate key areas where the US is doing well in reducing emissions and consumption and areas where we need to focus on improvement. We hope to assess the validity of the AEOs projections by comparing the future projections in the 2010 and 2015 report to the observed measures from the 2020 report. 
  - In order to look at what did the US do well in reducing emissions and consumption, we will compare our prediction of enery use in 2020 made from the 2010 and 2015 data to the actual data from 2020 to see whether the emissions and consumption decreased or increased. Also, we will look at individual energy source and sectors to see which of the reduced the most and which of them increased the most in the past decade and discuss the reason behind these changes.
  - 

### Background

Emissions Data: 
The emissions data measures outputs into the atmosphere in units of Million Metric Ton CO2 Equivalents. This is a measure of global warming potential. CO2 is a well know gas that traps heat in the atmosphere, however, there are other gasses which trap heat at a much higher magnitude. The measurement is called CO2 equivalents because it measures the amount of total global warming potential as a factor of the magnitude of one CO2 molecules warming potential. It is separated out by sector: residential, commercial, industrial, transportation, and electric power. As well as by fuel source: petroleum, natural gas, coal, and electric power generation, and an “other” category.  

Consumption Data: 
The consumption data measures how much energy was used in units of Quads. A Quad is a unit of energy equal to 10¹⁵ British Thermal Units, or 1.055 × 10¹⁸ joules in SI units. The US Department of Energy adopted the use of quads as their unit of energy use measurement for national and international reports.  The consumption data reports energy consumption by sector and power source. Sectors include residential, commercial, industrial, transportation, electric power, and delivered energy. The energy sources for consumption are more complex and vary with in each sector. 
 - We choose these data because they clearly presents the different meausres for emission and consumption of energy with labeled year, sector(the area that energy is being used) and source of energy. Thus, in order to see whether the US is doing well in reducing emissions and consumption and areas where we need to focus on improvement, we need to use these information to analyze and discuss about the energy performance in the US. 
 - The data was collected by the U.S Environmental Information Administration and is published on their website (https://www.eia.gov/outlooks/aeo/data) for public viewing. The projections were made using the AEO National Energy Modeling System. 
  - The energy used in the consumption data are measure by MmtCO2e (Million Metric Ton CO2) which is a common unit for measuring the energy emission and consumption. The energy used in the emission data are meausred by quads. These datasets divided the energy by its sector and source which makes the anlysis easier because we can look at different energy sector and source separately to see which of them contribute the most in the increasing of emission and consumption of energy and which of them reduced the most. Also, we choose to use the data within a 10-year period because not only the common energy source and how energy is used are similar within 10 year, but also because we can track the changes in different variables closely in a shorter period and time and compare the actual changes in 10 years with our prediction from the previous dataset.
    
    
### Analysis


## **(1) How do total emissions vary by year?**

```{r,echo = FALSE}

total_2011 = E2011A %>% 
    filter(Source == "Total" & Sector == "TotalbyFuel") %>% 
    arrange(Year)

sourcetopn_2011 = E2011A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
  filter(Year == 2010) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2015 = E2015A %>% 
    filter(Source == "Total") %>% 
    arrange(Year)

sourcetopn2015 = E2015A %>% 
    filter(!str_detect(Sector,"Total by Fuel")) %>% 
    filter(!str_detect(Source, "Total"))%>%
    filter(Year == 2014)%>% 
    slice_max(order_by = MmtCO2e, n = 1)

totalE2020 = E2020B %>% 
filter(Source == "Total") %>% 
arrange(Year)

sourcetopn2020 = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
  filter(Year == 2019) %>% 
    slice_max(order_by = MmtCO2e, n = 1)

```


```{r,echo = FALSE}
etotal2011 = ggplot(total_2011,aes(Year,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("2011 Data")+
    ylab(" Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5777, color = "red")

etotal2015 = ggplot(totalE2015, aes(YearNum,MmtCO2e)) +
    stat_smooth(se= FALSE) +
    geom_point()+
    ggtitle("2015 Data")+
    ylab(" Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 5499.3, color = "red")

etotal2020 = ggplot(totalE2020,aes(NumYear,MmtCO2e))+
    stat_smooth(se = FALSE)+
    geom_point()+
    ggtitle("2020 Data")+
    ylab("Total Emissions(MmtCO2e)")+
    xlab("Year")+
    ylim(4500,6500)+
    geom_vline(xintercept = 2020,  color = "red")+
    geom_hline(yintercept = 4993, color = "red")

grid.arrange(etotal2011,etotal2015,etotal2020, ncol = 3)

```

## **(2) Which fuel source is the greatest emissions contributer per year? How many million metric tons of CO2e are output by the greatest contributer?**

```{r,echo = FALSE}

etop2011 = ggplot(sourcetopn_2011, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2010")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)   

etop2015 = ggplot(sourcetopn2015, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2014")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)
  
    
etop2020 = ggplot(sourcetopn2020, aes(Source, MmtCO2e))+
    geom_col(fill = "blue")+
    ggtitle("2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Source")+
    ylim(0,2500)


grid.arrange(etop2011,etop2015,etop2020, ncol = 3)

```

## **(3)Which sector produces the most emissions? Does this vary by which year the outlook report is pulished?**

```{r,echo = FALSE}
sectormost2011A = E2011A %>% 
 filter(Source != "Total" & Sector != "TotalbyFuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
    
sectormost2015A = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>%
    filter(Sector != "Carbon Dioxide Emissions") %>%
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
sectormost2020A = E2020B %>% 
    filter(Source != "Total" & Sector != "Total by Fuel") %>% 
    group_by(Year, Sector) %>% 
    top_n(MmtCO2e, n = 1)
```


```{r,echo = FALSE}
sectormost2011B = E2011A %>% 
     filter(Source != "Total" & Sector != "TotalbyFuel", Year == 2009) %>% 
    group_by(Sector) %>% 
    slice_max(order_by = MmtCO2e, n = 1) %>% 
    arrange(desc(MmtCO2e)) %>% 
select(Sector, MmtCO2e) 


sectormost2015B = E2015A %>% 
    filter(Source != "Total" & Sector != "Total by Fuel", Year == 2013 & Sector != "Carbon Dioxide Emissions") %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector, MmtCO2e) 

sectormost2020B = E2020B %>% 
     filter(Source != "Total" & Sector != "Total by Fuel", Year == 2019) %>% 
    group_by(Sector) %>% 
    top_n(MmtCO2e, n = 1) %>%
        arrange(desc(MmtCO2e)) %>% 
    select(Sector,MmtCO2e) 
   


SM2011B = ggplot(sectormost2011B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
  xlab("Sector")+
    ylim(c(0,2200))
SM2015B = ggplot(sectormost2015B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
  xlab("Sector")+
    ylim(c(0,2200))
SM2020B = ggplot(sectormost2020B, aes(x = reorder(Sector, -MmtCO2e),y = MmtCO2e,fill = Sector))+
    geom_col()+
  xlab("Sector")+
    ylim(c(0,2200))

grid.arrange(SM2011B, SM2015B, SM2020B,nrow = 3)

```


## **(4)Looking at total emissions by fuel source, which is the greatest contributor? Which fuel source has the greatest projected consumption?**

```{r,echo = FALSE}
total_source_2011 <- E2011A %>%
    filter(!str_detect(Source, "Total"))%>%
    group_by(Source)%>%
    summarize(total_emission = sum(MmtCO2e)) %>%
    arrange(desc(total_emission))

BySourceTotal2010 <- E2011A %>%
    filter(Sector == "TotalbyFuel" & Source != "Total" & Year == 2010)

 

BySourceTotal2015 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total")

BySourceTotal2014 = E2015A %>% 
    filter(Sector == "Total by Fuel" & Source != "Total" & Year == 2014)

BySourceTotal2020 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Source != "Other")

BySourceTotal2019 = E2020B %>% 
    filter(Sector == "Total by Fuel"  & Source != "Total" & Year == 2019)

```

### Which fuel source is the greatest contribuer to emmisions?
```{r,echo = FALSE}

fuel2010 = ggplot(BySourceTotal2010, aes(x= Source, y = MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2010")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")
fuel2014 = ggplot(BySourceTotal2014, aes(Source,MmtCO2e, fill = Source))+
    geom_col() +
    ggtitle("Total Emmisions By Fuel Source: 2014")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

fuel2019 = ggplot(BySourceTotal2019, aes(Source,MmtCO2e, fill = Source))+
    geom_col()+
    ggtitle("Total Emmisions By Fuel Source: 2019")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Fuel Source")

grid.arrange(fuel2010,fuel2014,fuel2019)


```

##### Which fuel source has the greatest projected consumption?
```{r,echo = FALSE}

ggplot(BySourceTotal2020, aes(NumYear, MmtCO2e, color = Source))+
    geom_smooth(se = FALSE, method = "lm")+
    ggtitle("Total Emmisions Per Fuel Source: AEO 2020 Projections")+
    ylab("Emissions(MmtCO2e)")+
    xlab("Year")+
  theme(element_blank())
```

## **(5) Which sector uses the most renewables? Which uses the most non-renewables?**

```{r,echo = FALSE}

C2011Renew = C2011 %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable", Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>% 
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))

C2015Renew = C2015A %>%
    slice(1:3770) %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
                                        Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
C2020Renew = C2020B %>%
    mutate(RenewableResource = case_when(Source == "Renewable Energy" ~ "Renewable",
    Source != "Renewable Energy" ~ "Non-Renewable")) %>%
    filter(Sector == "Comericial" | Sector == "Industrial" | Sector == "Residential" | Sector == "Transportation") %>%
    group_by(Sector, RenewableResource) %>% 
    drop_na() %>% 
    summarise(energysum = sum(quads)) %>% 
    arrange(desc(energysum))
```

```{r,echo = FALSE}
renew_2011=ggplot(C2011Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", color = "lightblue")+
  facet_wrap(~RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
            subtitle = "Based on 2011 Data")+
     ylab("Consumptionr(quads)")+
    xlab("Sector")+
  ylim(0,3000)+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
    
renew_2015=ggplot(C2015Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", color = "blue") +
    facet_wrap(~ RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
            subtitle = "Based on 2015 Data")+
    ylab("Consumptionr(quads)")+
    xlab("Renewable")+
  ylim(0,3000)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
renew_2020 = ggplot(C2020Renew, aes(Sector, energysum))+
    geom_col(position = "dodge", color = "navy")+
    facet_wrap(~RenewableResource)+
    ggtitle("Which Sector Uses the Most Renewables?",
        subtitle = "Based on 2020 Data")+
    ylab("Consumptionr(quads)")+
    xlab("Renewable")+
  ylim(0,3000)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

grid.arrange(renew_2011,renew_2015,renew_2020, ncol =3)

```

## **(6) Which energy source is utilized the most in each sector?**
```{r echo=FALSE}
ratesector2010 = C2010B %>%
    filter(Year == 2009 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use and Related Statistics" & Sector != "Carbon Dioxide Emissions (million metric" ) %>%
    group_by(Sector, Year) %>%
    slice_max(order_by = GrowthRate, n = 1) %>%
    arrange(desc(GrowthRate)) %>%
    select(Sector, Source, GrowthRate,Year)%>%
    group_by(Source) %>%
    slice_max(order_by = GrowthRate, n = 1)
ratesector2015 = C2015B %>%
    slice(1:3770) %>%
    filter(Year == 2014 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    group_by(Sector, Year) %>% 
    slice_max(order_by = GrowthRate, n = 1) %>% 
    arrange(desc(GrowthRate)) %>%
    select(Sector, Source, GrowthRate,Year)%>%
    group_by(Source) %>%
    slice_max(order_by = GrowthRate, n = 1)
ratesector2020 = C2020B %>%
    filter(Year == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    group_by(Sector, Year) %>% 
    slice_max(order_by = GrowthRate, n = 1) %>% 
    arrange(desc(GrowthRate)) %>% 
    select(Sector,Source,GrowthRate,Year)%>%
    group_by(Source) %>%
    slice_max(order_by = GrowthRate, n = 1)
ratesector = bind_rows(ratesector2010,ratesector2015,ratesector2020)

```

## **(7)Which sector has the highest growth rate for each fuel source?**

```{r, include=FALSE}
ratesource2010 = C2010B %>%
    filter(Year == "2010" & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use and Related Statistics") %>% 
    select(Source,Sector,GrowthRate)%>%
    arrange(desc(GrowthRate))%>%
    slice_max(order_by = GrowthRate, n = 1) 

ratesource2015 = C2015B %>%
    filter(Year == 2014 & Sector != "NA" & 
               Sector != "Unspecified Sector" & 
               Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate)

ratesource2020 = C2020B %>%
    filter(Year == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate)
bind_rows(ratesource2010,ratesource2015,ratesource2020)

```

```{r,echo = FALSE}
ggplot(ratesector, aes(Source, GrowthRate, fill= Sector) ) +
    geom_col()+
    facet_wrap( ~ Year, nrow = 1) + 
    ggtitle("Which sector has the highest growth rate for each fuel source?")+
    ylab("Growth Rate(%)")+
    xlab("Source")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## **(8) Which energy source has the highest growth rate over all?**

```{r,echo = FALSE}
ratesource2011 = C2011 %>%
    filter(Year == 2010 & Sector != "NA") %>% 
    filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" | Sector == "Comericial" | Sector == "Delivered Energy Consumption" | Sector == "Electric Power ") %>% 
    select(Source,Sector,GrowthRate,Year)%>%
    slice_max(order_by = GrowthRate, n = 1)

ratesource2015 = C2015B %>%
    filter(Year == 2014 & Sector != "NA" & 
               Sector != "Unspecified Sector" & 
               Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate,Year)

ratesource2020 = C2020B %>%
    filter(Year == 2019 & Sector != "NA" & Sector != "Unspecified Sector"& Sector != "Energy Use & Related Statistics") %>% 
    slice_max(order_by = GrowthRate, n = 1) %>%
    select(Source,Sector,GrowthRate,Year)

highgrowthrate = bind_rows(ratesource2011,ratesource2015,ratesource2020)

ggplot(highgrowthrate,aes(Year,GrowthRate, fill = Source))+
    geom_col()+
     theme(
  # Hide panel borders and remove grid lines
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  axis.line = element_line(colour = "black"))+
    annotate("text", x=c(2010,2014,2019), y=5, label = c("2010","2014","2019"), color = "black", size = 5)+
  scale_x_discrete()

```

## **(9)How do energy consumption projection trends change based on the year the report was published?**

```{r,echo = FALSE}
ConsumptionByYear_2011 = C2011 %>% 
    filter(Source == "Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
ConsumptionByYear2015 = C2015B %>% 
    filter(Source == "Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
ConsumptionByYear2020 = C2020B %>% 
    filter(Source =="Total") %>% 
    group_by(Year) %>% 
    summarise(sumquads = sum(quads))
    
q82011 = ggplot(ConsumptionByYear_2011, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2011 Data") +
    ylab("Energy Consumption(quads)") +
    xlab("Year")+
    xlim(2020,2050)+
    ylim(225,375)+
     theme_dark()
q82015 = ggplot(ConsumptionByYear2015, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2015 Data") +
    ylab("Energy Consumption(quads)") +
    xlab("Year")+
    xlim(2020,2050)+
    ylim(225,375)+
     theme_dark()
q82020 = ggplot(ConsumptionByYear2020, aes(Year, sumquads))+
    geom_line(color  = "red", size = 1) +
    ggtitle("Total Fuel Consumtion Projections", 
            subtitle = "Based on 2020 Data") +
    ylab("Energy Consumption(quads)") +
    xlab("Year")+
      xlim(2020,2050)+
    ylim(225,375)+
    theme_dark()

grid.arrange(q82011,q82015,q82020)
   
    
    
```

## **(10).Compare projections for 2019 agaisnt to actual measures in 2019.**

```{r,echo = FALSE}
Actual2019 = C2020B %>% 
    filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ") %>%
  mutate(yearFrom = "2019")

Prediction2015 = C2015B %>% 
        filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ") %>%
  mutate(yearFrom = "2015")

Prediction2011 = C2011 %>% 
       filter(Year == 2019) %>% 
  filter(Sector == "Residential" | Sector == "Industrial" | Sector == "Transportation" |  Sector == "Electric Power ") %>%
  mutate(yearFrom = "2011")

predictions_actual <- bind_rows(Actual2019, Prediction2015, Prediction2011) %>%
  group_by(yearFrom)

ggplot(predictions_actual, aes(x = Sector, y = quads, fill = yearFrom)) +
  geom_col(position = "dodge", alpha = 0.5)+
  ggtitle("Actual Values For Year 2019", subtitle ="From AEO 2020")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,50)
```
```{r}
ggplot(Actual2019, aes(x = reorder(Sector, -quads), quads))+
    geom_col(fill = "lightblue")+
  ggtitle("Actual Values For Year 2019", subtitle ="From AEO 2020")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)
  
ggplot(Prediction2011, aes(x = reorder(Sector, -quads), quads))+
geom_col(fill = "blue")+
  ggtitle("Predicted Values For Year 2019", subtitle =  "From AEO 2011")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)
    
ggplot(Prediction2015, aes(x = reorder(Sector, -quads), quads))+    
geom_col(fill = "navy")+
  ggtitle("Predicted Values For Year 2019", subtitle =  "From AEO 2015")+
  ylab("Energy Consumption(Quads)")+
  xlab("Sector")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylim(0,120)

```

```{r}
ConsumptionByYear2015 <- ConsumptionByYear2015 %>%
    mutate(Dataset = "2015")
ConsumptionByYear2020 <- ConsumptionByYear2020 %>%
    mutate(Dataset = "2020")
consumptions_2015_2020_combined <- bind_rows(ConsumptionByYear2015, ConsumptionByYear2020)
sourcetopn2015 <- sourcetopn2015 %>%
    mutate(Dataset = "2015")
sourcetopn2020 <- sourcetopn2020 %>%
    mutate(Dataset = "2020")
emission_2015_2020_combined <- bind_rows(sourcetopn2015, sourcetopn2020)

ggplot(consumptions_2015_2020_combined, aes(Year, sumquads, fill = Dataset)) +
geom_col()
    
ggplot(emission_2015_2020_combined, aes(Year, MmtCO2e, fill = Dataset)) +
    geom_col()
```

#### Compare linear models for 2020-2030 time span between the 2015 and 2020 outlook data. Hypothesis test for different slopes. 
```{r}

    
```

## 12.How does the use of Petroleum and Coal changed from 2010-2019 period to 2020-2030 period?
### Petroleum

 - From the previous table, we can see that Petroleum is a main source of energy contributing to the energy emission. As the technology and use of energy changed and developed throughout the decase, we are interested in how the use of petroleum changes in the sector of transportation.  From the previous problem, we can see that the main sector uses petroleum is in transportation. We now want to compare the result within 20 dacades to see whether the mean use of Coal for emission changes significantly.

```{r echo=FALSE}
p_c2010 = E2010A%>%
    filter(Source == "Petroleum" & Sector != "Total by Fuel")%>%
    filter(Year< 2020)
p_c2020 = E2020B%>%
    filter(Source == "Petroleum" & Sector != "Total by Fuel")%>%
    filter(Year < 2030 & Year >= 2020)
p_c = bind_rows(p_c2010,p_c2020)%>%
    filter(Sector == "Transportation")%>%
    mutate(range = case_when(Year < 2019 ~ "2010",
                     Year >= 2019 ~ "2020"))
sum_PC = p_c%>%
    group_by(range)%>%
    summarize(n = n(),
              sum_MmtCO2e = sum(MmtCO2e),
              mean = mean(sum_MmtCO2e),
              sd = sd(MmtCO2e))
ggplot(p_c, aes(x=log10(MmtCO2e),color=range)) +
  geom_density() +
  xlab("MmtCO2e(log10)") +
  ylab("Density") +
  ggtitle("Mean source of energy in Emission (MmtCO2e)",
          subtitle="early = 2010-2020, recent = 2020-2030") +
  theme_bw()
```
- Thus, We are 95% confident that the mean emission energy from Petroleum meausred by MmtCO2e in period 2009-2018 was between 227.9347 MmtCO2e lower to 296.1169 MmtCO2e higher than the mean emission energy from Petroleum meausred by MmtCO2e in period 2019-2029.

 - Now, using the mean and se generated from the previous t-test, we now are able to verify the p value by calculating the test statstistics.
```{r echo=FALSE}
tstat_p <- mean_diff_p/se_p
## pvalue
pt(-abs(tstat_p), dof_p) + pt(abs(tstat_p), dof_p, lower.tail=FALSE)
```
 - The p value is 0.0002402814 (less than 0.05 (typically ≤ 0.05)) is statistically significant. It indicates strong evidence against the null hypothesis, as there is less than a 5% probability the null is correct (and the results are random). Therefore, we reject the null hypothesis, and accept the alternative hypothesis of the mean use of petroleum in transportation is different from 2010-2020 period to 2020-2030 period. 
### Coal

- From the previous table, we can see that Coal is also a main source of energy contributing to the energy emission. Also, we noticed that the major use of Coal is in Sector of Electric Power. As the technology and use of energy changed and developed throughout the decase, we are instrerested in how the use of Coal changes in the sector of electric power. We now want to compare the result within 20 dacades to see whether the mean use of Coal for emission in Electric Power changes significantly.
```{r echo=FALSE}
c2010 = E2010A%>%
    filter(Source == "Coal" & Sector != "Total by Fuel")%>%
    filter(Sector == "Electric Power")%>%
    filter(Year < 2019)
c2020 = E2020B%>%
    filter(Source == "Coal" & Sector != "Total by Fuel")%>%
    filter(Sector == "Electric Power")%>%
    filter(Year < 2030 & Year >=2020)
c = bind_rows(c2010,c2020)%>%
    mutate(range = case_when(Year < 2019 ~ "2010",
                     Year >= 2019 ~ "2020"))
sum_C = c%>%
    group_by(range)%>%
    summarize(n = n(),
              sum_MmtCO2e = sum(MmtCO2e),
              mean = mean(sum_MmtCO2e),
              sd = sd(MmtCO2e))
ggplot(c, aes(x=log10(MmtCO2e),color=range)) +
  geom_density() +
  xlab("MmtCO2e(log10)") +
  ylab("") +
  ggtitle("Coal Use in Electric Power for Energy Emission",
          subtitle="early = 2010-2020, recent = 2020-2030") +
  theme_bw()
```
#### Test the hypothesis that the mean energy use from Coal in 2010-2020 (Pa) is equal to that in 2020-2030 period (Pb) versus that alternative that they are different. 

$$
H_0: P_a = P_b
$$

$$
H_a:  P_a \neq P_b
$$
```{r echo=FALSE}
x1_c <- c %>%
  filter(range == "2010") %>%
  pull(MmtCO2e)
x2_c <- c %>%
  filter(range == "2020") %>%
  pull(MmtCO2e)
```
```{r echo=FALSE}
t.test(x1_c,x2_c)
mean_diff_c <- mean(x1_c) - mean(x2_c)
dof_c <- 16.803 ## from t.test() above
se_c <- sqrt(sd(x1_c)^2/length(x1_c) + sd(x2_c)^2/length(x2_c))
ci_5_c <- mean_diff_c + c(-1,1)*qt(0.975,dof_c)*se_c
signif(ci_5_c,3)
```
We are 95% confident that the mean emission energy from Coal meausred by MmtCO2e in period 2009-2018 was between 37.6 MmtCO2e lower to 501.0 MmtCO2e higher than the mean emission energy from Coal meausred by MmtCO2e in period 2019-2029.


Test the hypothesis that the mean energy use from Coal in 2010-2019 period is equal to that in the recent era versus that alternative that they are different. 

```{r echo=FALSE}
tstat_c <- mean_diff_c/se_c
## pvalue
pt(-abs(tstat_c), dof_c) + pt(abs(tstat_c), dof_c, lower.tail=FALSE)
```
 - The p value is 4.151938e-51 (less than 0.05 (typically ≤ 0.05)) is statistically significant. It indicates strong evidence against the null hypothesis, as there is less than a 5% probability the null is correct (and the results are random). Therefore, we reject the null hypothesis, and accept the alternative hypothesis of the mean use of Coal in electric power is different from 2010-2020 period to 2020-2030 period. As renewable and other more efficient energy source being used, from the graph in previous question, we can seethe use of coal decline in electric power which contributing much lesser in energy emission. This is a good news because the coal-fired power plants release more greenhouse gases per unit of energy produced than any other electricity source (according to https://www.greenamerica.org/fight-dirty-energy/amazon-build-cleaner-cloud/coal-why-it-dirty).

#### Test the hypothesis that the mean energy use from Petroleum in 2010-2020 (Pa) is equal to that in 2020-2030 period (Pb) versus that alternative that they are different. 

$$
H_0: P_a = P_b
$$

$$
H_a:  P_a \neq P_b
$$
```{r echo=FALSE}
x1_p <- p_c %>%
  filter(range == "2010") %>%
  pull(MmtCO2e)
x2_p <- p_c %>%
  filter(range == "2020") %>%
  pull(MmtCO2e)

t.test(x1_p,x2_p)
mean_diff_p <- mean(x1_p) - mean(x2_p)
dof_p <- 15.664 ## from t.test() above
se_p <- sqrt(sd(x1_p)^2/length(x1_p) + sd(x2_p)^2/length(x2_p))
ci_5_p <- mean_diff_p + c(-1,1)*qt(0.975,dof_p)*se_p
```

### Discussion

Param:
Broadly speaking, based on the 2015 and 2020 datasets, it is certain that emissions have dropped. However, it is also evident that consumption will either mostly stay steady or rise slowly. Projections for consumption skyrocket in each dataset regardless of year, so it does appear the AEO is taking the pessimistic outlook for their projections. 

The datasets for 2015 energy outlook seems to be slightly inconsistent with format and depth of projections with other datasets. This is not too much of a shortcoming, but as analysts, it does leave us guessing for the verification of some of the prior projections in later datasets within a single dataset. Additionally, the outlook is made during the fiscal year that the outlook is released, so only the previous few years will be accurate. This requires more data to be obtained, which in turn requires more standardization between the format of each dataset. Perhaps, given the previous years data-sets and actual values, a match between the mesh of datas that are equal could come handy. And to test performance even further, a whole parity chart could be made between different datasets actual performance and projected performance.

Further analysis can be made on specific governmental body's policy performance, specifically around the time period they are made in tandem to the associated energy outlook data sets. Similarly, new data can be collected regarding this.

Conclusively, it appears Americans are doing a very good job on curbing emissions when considering the 2015 and 2020 dataset. Consumption is still there, but it is highly unlikely an increasing population will decrease their consumption of energy unless several billions of technology's source of energy is changed to a more modern source, this also requiring a change in the technology itself.

Yunqing:
- for data in 2011 delivered energy consumptions contribute the most among all sectors of energy and coal is the most commonly used source of energy followed by electricity and petroleium. Transportation and electric power produces the most parts of the energy.And a huge portion of the energy is used for US GDP. Through these comparision, it will be interesting to look at how it changed trhoughout the decade.
- one of the potential short-comings of this analysis will be the sized of our datasets are slightly different which might caused a little inaccuracy in final analysis
- we can look more deeply into how different variables changes throughout the decades and discuss the potential caused and effect of the changes such as how the a environmental related policy or crisis impacted the energy use or how some major changes in economy or manufacturing market affect the use of energy
 - my new questions is can we use some of the statistc model to predict the future changes based on previous year's dataset which we can later use to compare with the actual situation and talk about why there are potential difference between prediction and the reality
 - Instead of comparing all sectors and sources, I will like to look closely at a specific source and sector and study it more in-depth
- I will like to look at some data that related to the variable which changed the most throughout the decade and study the reason behind the change
- In conclusion, based on the previous analysis, I would like to predict that the consumption of energy continue to increase in the next decade but other source of the energy will surpase the domination of energy use of coal becasue we can see the curve of electric power increasing with a higher slope the coal in 2011
