---
title: 'Project Proposal Group 6: 325'
author: "Lindsey Fisher"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Project Proposal

### The names of the students in the group.
Lindsey Fisher
Param Bhandare
YUNQING XIAO

### A brief description of how the group will communicate and work together.
After discussion on Thursday at 10:45 am CT

### What platforms will the group use to communicate and share documents.
Google Documents, UW-Madison Zoom 

### The question or questions of interest that motivate the planned study.

#### 1. Carbon Dioxide Emissions by Sector and Source (DATA SET 1)
>How does total emissions vary by each year? By sector?
>Which sector produces the most emissions? Does this vary by year?
>Looking at total emissions by fuel? Which is the greatest contributor in 2011? Which has the greatest projected consumption in 2020?

#### 2.Energy Consumption by Sector and Source  (DATA SET 2)
>Which sector uses the most renewables? Which uses the most non-renewables?
>Which energy source is utilized the most in each sector?
>Which sector has the highest growth rate for each energy source?
>Which energy source has the highest growth rate over all? 
>How does energy consumption vary by year?

#### 3.Compare projections in 2011 to actual measures in 2020(DATA SET 3 and 4)

```{r}
energyemmisions2011 = read_csv("emissions2011.csv")
head(energyemmisions2011)

energyconsumption2011 = read_csv("consumption2011.csv")
head(energyconsumption2011)

energyemissions2020 = read_csv("emissions2020.csv")
head(energyemissions2020)

energyconsumption2020 = read_csv("consumption2020.csv")
head(energyconsumption2020)

```

### Data sources: 



>> Annual Energy Outlook 2020: Table 18. Energy-Related Carbon Dioxide Emissions by Sector and Source
>> Annual Energy Outlook 2020: Table: Table 2. Energy Consumption by Sector and Source
 
 https://www.eia.gov/outlooks/aeo/data/browser/#/?id=17-AEO2020&cases=ref2020&sourcekey=0
 

>> Annual Energy Outlook 2011: Carbon Dioxide Emissions by Sector and Source

https://openei.org/doe-opendata/dataset/140e63e5-f215-4e15-a0e0-b1784778ff0c

>> Annual Energy Outlook 2011: Energy Consumption by Sector and Source, United States

https://openei.org/doe-opendata/dataset/4036890e-e139-4759-87ef-7814e814234e


### Methods 

>> We can use scatter plots or bar graph to show the relationship between each variable and the change of years
>> We can use linear regression model to see the trend of data throughout the year
>> We can use various join method to combine different data together to see how they affect each other

### Possible model 

>> We can build a linear model or other statistical distribution model to find the relationship between each variable
>> Based on the models we build we can compare the actual results in the model with our predictions
